---
title: "English Language Throughout Centuries"
author: "Grzegorz Krochmal & Katarzyna Kryńska"
date: "4 01 2020"
output: ioslides_presentation
      # incremental: true
---


```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)

setwd("C:/Users/grzeg/Desktop/studia/Data Science/2 rok/semestr 1/Advanced_VisualisationR/projekt/Adv_Vis_R_Project")
'17th_data' <- data.frame(readxl::read_xlsx("17th_joined_file.xlsx"))
'18th_data' <- data.frame(readxl::read_xlsx("18th_joined_file.xlsx"))
'19th_data' <- data.frame(readxl::read_xlsx("19th_joined_file.xlsx"))
'20th_data' <- data.frame(readxl::read_xlsx("20th_joined_file.xlsx"))
authors_data <- data.frame(readxl::read_xlsx("Author_birthplace_with_geo_data.xlsx"))
```

## Goals of the project

> - See the evolution of English language in literature between 17th - 20th century
> - Cross Century Analysis of Sentiment
> - Cross Century Analysis of Parts of Speech

## Dataset Description

> - Books collected from Guttenberg Project
> - Sample of books for centuries based on rank of most popular ones from www.goodreads.com

## Dataset Preparation

> - Only words extracted from books with regex
> - Lemmatization introduced with Stanford NLP Library - Python
> - Sentiment Analysis of Lemmatized Words with Vader - Python



# Initial analysis

## Number of Authors vs Number of Titles

```{r echo = FALSE}
centuries <- c("17th", "18th", "19th", "20th")
number_of_titles <- c(15, 15, 15, 14)

century_basic <- data.frame(centuries, number_of_titles)
century_basic <- cbind(century_basic, authors_data %>% group_by(Century) %>% tally(name = "Number of Authors") %>% select_at("Number of Authors"))
century_basic_long <- melt(century_basic, id=c("centuries"))

ggplot(century_basic_long) + geom_bar(aes(x = centuries, y = value, fill = variable), stat = "identity", position = "dodge", width = 0.7) + 
  scale_fill_manual("Variable\n", values = c("orange", "blue"), labels = c("Number of Titles", "Number of Authors")) + 
  labs(x = "\nCenturies", y = "\nNumber of") + 
  theme_bw(base_size = 14) 
  
```

## Total Number of Words vs Count of Unique Words for each Century

```{r echo = FALSE}
num_of_occ <- c(sum(`17th_data`$Number.of.occurences), sum(`18th_data`$Number.of.occurences), sum(`19th_data`$Number.of.occurences), sum(`20th_data`$Number.of.occurences))

num_of_unique <- c(length(unique(`17th_data`$Lemma)), length(unique(`18th_data`$Lemma)), length(unique(`19th_data`$Lemma)), length(unique(`20th_data`$Lemma)))

century_basic <- cbind(century_basic, num_of_occ, num_of_unique)

century_basic_long <- melt(century_basic[,c("centuries", "num_of_occ", "num_of_unique")], id=c("centuries"))

ggplot(century_basic_long) + geom_bar(aes(x = centuries, y = value, fill = variable), stat = "identity", position = "dodge", width = 0.7) + 
  scale_fill_manual("Variable\n", values = c("orange", "blue"), labels = c("Count of All Words", "Count of Unique Words")) + 
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +
  labs(x = "\nCenturies", y = "\nNumber of") + 
  theme_bw(base_size = 14)


```

## Count of Unique Words to All Words

```{r echo = FALSE}
century_basic <- century_basic %>% mutate(Unique_to_all = round(num_of_unique/num_of_occ,4))

ggplot(century_basic) + geom_bar(aes(x = centuries, y = Unique_to_all), stat = "identity", width = 0.7, fill ="orange") + 
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +
  labs(x = "\nCenturies", y = "\nProportion") + 
  theme_bw(base_size = 14)
```


# Part of speech analysis

```{r echo = FALSE}
`17th_data`$Part_of_speech <- ifelse(`17th_data`$Part_of_speech %in% c("VERB", "ADJ", "ADV", "NOUN"), `17th_data`$Part_of_speech, "OTHER")
`18th_data`$Part_of_speech <- ifelse(`18th_data`$Part_of_speech %in% c("VERB", "ADJ", "ADV", "NOUN"), `18th_data`$Part_of_speech, "OTHER")
`19th_data`$Part_of_speech <- ifelse(`19th_data`$Part_of_speech %in% c("VERB", "ADJ", "ADV", "NOUN"), `19th_data`$Part_of_speech, "OTHER")
`20th_data`$Part_of_speech <- ifelse(`20th_data`$Part_of_speech %in% c("VERB", "ADJ", "ADV", "NOUN"), `20th_data`$Part_of_speech, "OTHER")
```


## Percentage Pie Chart of Parts of Speech for Unique Words

```{r echo = FALSE}
pie_chart <- function(df, title){
  pie_data <- df %>% group_by(Part_of_speech) %>% summarise(Percentage = round(n()/length(df$Part_of_speech) * 100,2))
  pie_data <- pie_data %>% mutate(lbls = paste(Part_of_speech, Percentage, "%"))
  pie(pie_data$Percentage, labels = pie_data$lbls, col = rainbow(length(pie_data$lbls)), 
      main = title)
}

par(mfrow=c(2,2), oma = c(0, 0, 2, 0), mai = c(0, 0.1, 0.5, 0.1))
pie_chart(`17th_data`, "17th Century")
pie_chart(`18th_data`, "18th Century")
pie_chart(`19th_data`, "19th Century")
pie_chart(`20th_data`, "20th Century")
```


## Stacked Barplot with Percentage of Parts of Speech for Unique Words

```{r echo = FALSE}
Part_of_speech <- cbind(centuries = c(rep(c("17th"),5), rep(c("18th"),5), rep(c("19th"),5), rep(c("20th"),5)),
  rbind(data.frame(`17th_data` %>% group_by(Part_of_speech) %>% summarise(Percentage = round(n()/length(`17th_data`$Part_of_speech) * 100,2))),
      data.frame(`18th_data` %>% group_by(Part_of_speech) %>% summarise(Percentage = round(n()/length(`18th_data`$Part_of_speech) * 100,2))),
      data.frame(`19th_data` %>% group_by(Part_of_speech) %>% summarise(Percentage = round(n()/length(`19th_data`$Part_of_speech) * 100,2))),
      data.frame(`20th_data` %>% group_by(Part_of_speech) %>% summarise(Percentage = round(n()/length(`20th_data`$Part_of_speech) * 100,2)))))


ggplot(Part_of_speech, aes(x = centuries, y = Percentage, fill = Part_of_speech)) + geom_bar(stat = "identity", position = "stack", width = 0.7) + 
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +
  geom_text(aes(label = paste0(Percentage,"%")), 
            position = position_stack(vjust = 0.5), size = 3) +
  labs(x = "\nCenturies", y = "\nPercentage", fill = "Part of Speech") +
  theme_bw(base_size = 14)   
```

# Sentiment Analysis


## Sentiment Type to Sum of All Words
```{r echo = FALSE, message = FALSE}
senti_data <- rbind(
                    data.frame(Centuries = "17th", 
                             neg = sum(`17th_data`$neg)/length(`17th_data`$neg), 
                             neu = sum(`17th_data`$neu)/length(`17th_data`$neu),
                             pos = sum(`17th_data`$pos)/length(`17th_data`$pos)),
                    data.frame(Centuries = "18th", 
                             neg = sum(`18th_data`$neg)/length(`18th_data`$neg), 
                             neu = sum(`18th_data`$neu)/length(`18th_data`$neu),
                             pos = sum(`18th_data`$pos)/length(`18th_data`$pos)),
                    data.frame(Centuries = "19th", 
                             neg = sum(`19th_data`$neg)/length(`19th_data`$neg), 
                             neu = sum(`19th_data`$neu)/length(`19th_data`$neu),
                             pos = sum(`19th_data`$pos)/length(`19th_data`$pos)),
                    data.frame(Centuries = "20th", 
                             neg = sum(`20th_data`$neg)/length(`20th_data`$neg), 
                             neu = sum(`20th_data`$neu)/length(`20th_data`$neu),
                             pos = sum(`20th_data`$pos)/length(`20th_data`$pos))
                    )

melt(senti_data) %>% ggplot(aes(x = Centuries, y = value, fill = variable)) + geom_bar(stat = "identity", position = "stack", width = 0.7) + 
  scale_y_continuous(labels = function(x) paste0(x * 100, "%")) +
  geom_text(aes(label = paste0(round(value * 100,2),"%")), 
           position = position_stack(vjust = 0.5), size = 3) + 
  labs(fill = "Sentiment", y = "Percentage") +
  theme_bw(base_size = 14)
```